---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# causalgraphs

<!-- badges: start -->
<!-- badges: end -->

**causalgraphs** provides fast, flexible, and cross-language causal graph (DAG) manipulation in R, powered by Rust via [extendr](https://extendr.github.io/).  
It is ideal for scientific computing, causal inference, and graph-based modeling.

## Features

- Add/remove nodes and edges in directed acyclic graphs (DAGs)
- Mark nodes as latent (unobserved)
- Query parents, children, ancestors, and other graph properties
- Efficient Rust backend for high performance
- Seamless R interface

## Installation

First, ensure you have a working Rust toolchain. You can install it by following the instructions at rustup.rs.
You can install the development version of `causalgraphs` from GitHub. We recommend using the `pak` package for a smooth installation of all dependencies.

```r
# install.packages("pak")
pak::pkg_install("pgmpy/causalgraphs/r_bindings/causalgraphs")

# Or using devtools:
# install.packages("devtools")
devtools::install_github("pgmpy/causalgraphs", subdir = "r_bindings/causalgraphs")
```

## Development

- Rust source: `src/rust/`
- R wrappers: `R/`
- Test and experiment: `test.R`

If you want to contribute to the package, you'll need to build it from a local clone of the repository.

**Clone the repository:**
```sh
  git clone https://github.com/pgmpy/causalgraphs.git
  cd causalgraphs/r_bindings/causalgraphs
```

**Build and Test in R:**
From an R session inside the `r_bindings/causalgraphs` directory:
```r
# Sync R ↔ Rust bindings:
# 1. Compiles the Rust crate into a shared library (causalgraphs.so/.dll/.dylib)
# 2. Generates/updates R wrapper functions in R/extendr-wrappers.R
# Run this whenever you change the Rust code or add new #[extendr] functions.
rextendr::document()

# Load the package into R
devtools::load_all()
```

## Developer Notes

Whenever you're working on both `rust_core` and `r_causalgraphs` locally, make sure you comment the github `rust_core` dependency in `Cargo.toml` and uncomment the local path dependency if you are developing locally.

## Example

```{r example, eval=FALSE}
library(causalgraphs)

# Create a new DAG
dag <- RDAG$new()

# Add nodes
dag$add_node("A", latent = FALSE)
dag$add_node("B", latent = FALSE)
dag$add_node("L", latent = TRUE)  # Latent node

# Add edges
dag$add_edge("A", "B", 10)
dag$add_edge("B", "C", 20)

# Inspect graph
cat("Nodes:", dag$nodes(), "\n")
cat("Latents:", dag$latents(), "\n")
cat("Node count:", dag$node_count(), "\n")
cat("Edge count:", dag$edge_count(), "\n")

# Get edges
edges <- dag$edges()
cat("Edges:\n")
print(edges)

# Query relationships
cat("Parents of B:", dag$get_parents("B"), "\n")
cat("Children of A:", dag$get_children("A"), "\n")
cat("Ancestors of C:", dag$get_ancestors_of(c("C")), "\n")
```

## License

MIT

---

*This README was generated from `README.Rmd`. Please edit that file and knit to update.*